DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-17 paper/evaluation.tex
revision up to Section 4.E

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact on their inference
timings. In other words, we are interested in how shared resource
contention affects inference timing. These instances are not identical
copies of the same model, but are instead copies of different models. 
This is done to ensure that no L2 cache memory is shared between the 
models as they are running.

@@

These increases in inference times in the co-scheduled scenarios are
expected and are caused by contention in the shared hardware
resources, such as the shared L2 cache and/or the DRAM controller.

@@

\subsection{Effect of Co-scheduling Memory Performance Hogs}\label{sec:eval-memhog}

In this experiment, we investigate the impact of contended DRAM
requests to the DNN inference timing of DeepPicar.
For the experiment, we use a synthetic memory
intensive benchmark from the IsolBench suite~\cite{Valsan2016}.
We run a single DNN model on one core, and
co-schedule an increasing number of the memory intensive synthetic
benchmarks~\footnote{We use the \emph{Bandwidth} benchmark in the
  IsolBench suite, with the following command line parameters: \texttt{\$
  bandwdith -a write -m 16384 -t 1000}}, on the remaining idle cores.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
  \caption{Effect of memory performance hogs on the shared resources. 
    The DNN model uses Core 0 and memory-hog co-runners
    use the rest of the cores.}
  \label{fig:perf_vs_bandwidth_mem}
\end{figure}

Figure~\ref{fig:perf_vs_bandwidth} shows the execution time of the DNN 
model and Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
total memory bus accesses (first group) and L2 miss rates (second group).
In both figures, the DNN model is running on Core 0 as a function of the 
number of co-scheduled memory intensive synthetic benchmarks.
%% We also show 
%% the normalized total system bus access count (middle group), which is
%% representative of memory bandwidth usage of all the cores.
First, as we
increase the number of co-runners, the DNN model's execution times are
increased exponentially---by up to 11.6X---even though the DNN model
is running on a dedicated core (Core 0). On the other hand, the DNN
model's L2 cache-miss rates do not increase as much.
In other words, the DNN model's execution increase is not fully
explained by increased contention in the L2 cache space.
Instead, the exponential increase in the total number of bus accesses,
due to the memory-hog co-runners, appears to be more closely
correlated with the DNN model's execution time increase. This means
the high memory bandwidth pressure from the co-scheduled
memory intensive benchmarks increased the DNN model's
memory access latency.

@@

This observation suggests that shared cache partitioning
techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
solutions for DeepPicar's DNN processing workload, as cache
partitioning does not provide memory performance guarantees.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
  \caption{DNN model performance on different in-order
ARM Cortex cores.}
  \label{fig:a53_vs_a7}
\end{figure}

We do find the performance of the model in the presence of 
write co-runners surprising, as it has been previously shown in experiments
on another in-order ARM core, the Cortex A7, that write interference was not as 
significant\cite{Valsan2016}. We also tested the model, in the presence of write 
co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
As expected, when the model is run on A7 cores, the intereference of the write
co-runners is not as impactful on model performance as it can still operate at 
under 10 Hz.

>>>>

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact on their inference
timings. In other words, we are interested in how shared resource
contention affects inference timing. These instances are not identical
copies of the same model, but are instead copies of different models
(but of the same network architecture). 
This is done to ensure that no L2 cache memory is shared between the 
models as they are running.

@@

These increases in inference times in the co-scheduled scenarios are
expected due to contention in the shared hardware
resources, such as the shared cache and DRAM.

@@

\subsection{Effect of Co-scheduling Synthetic Benchmarks}\label{sec:eval-memhog}

In this experiment, we further investigate the impact of shared
resource contention to the DNN inference timing of DeepPicar using  
synthetic memory benchmarks. Specifically, we use the \emph{Bandwidth}
benchmark from the IsolBench suite~\cite{Valsan2016}, which
sequentially reads or writes a big array; we henceforth refer BwRead 
as Bandwidth with read accesses and BwWrite as the one with write
accesses. The experiment setup is as follows. We run a single DNN
model on one core, and co-schedule an increasing number of the
Bandwidth benchmark instances on the other cores. We repeat the
experiement first with BwRead~\footnote{\texttt{\$ bandwdith -a read
    -m 16384 -t 1000}} and next with BwWrite~\footnote{\texttt{\$
    bandwdith -a write -m 16384 -t 1000}}.

@@

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth_mem}
%%   \caption{Effect of memory performance hogs on the shared resources. 
%%     The DNN model uses Core 0 and memory-hog co-runners
%%     use the rest of the cores.}
%%   \label{fig:perf_vs_bandwidth_mem}
%% \end{figure}

Figure~\ref{fig:perf_vs_bandwidth} shows the execution times of the DNN 
model against varying number of BwRead and BwWrite co-runners. First,
note that BwWrite cause significantly high execution time incrase---up
to 11.6X---on the DNN inferencing timing while BwRead causes
relatively much smaller time increase. While execution time increase
was expected, the degree of which is quite surprising---\emph{if the DNN
controller was driving an actual car, it would result in crash of the
car!}

\fixme{story: is cache space contention the reason? traditional
  solution is cache partitioning. so, next we evaluate if cache
  partitioning can solve the problem. spolier alert: it does not.}

%% Figure~\ref{fig:perf_vs_bandwidth_mem} shows both the normalized
%% total memory bus accesses (first group) and L2 miss rates (second group).
%% In both figures, the DNN model is running on Core 0 as a function of the 
%% number of co-scheduled memory intensive synthetic benchmarks.
%% %% We also show 
%% %% the normalized total system bus access count (middle group), which is
%% %% representative of memory bandwidth usage of all the cores.
%% First, as we
%% increase the number of co-runners, the DNN model's execution times are
%% increased exponentially---by up to 11.6X---even though the DNN model
%% is running on a dedicated core (Core 0). On the other hand, the DNN
%% model's L2 cache-miss rates do not increase as much.
%% In other words, the DNN model's execution increase is not fully
%% explained by increased contention in the L2 cache space.
%% Instead, the exponential increase in the total number of bus accesses,
%% due to the memory-hog co-runners, appears to be more closely
%% correlated with the DNN model's execution time increase. This means
%% the high memory bandwidth pressure from the co-scheduled
%% memory intensive benchmarks increased the DNN model's
%% memory access latency.

@@

%% This observation suggests that shared cache partitioning
%% techniques~\cite{Gracioli2015,Kim2016} may not be effective isolation
%% solutions for DeepPicar's DNN processing workload, as cache
%% partitioning does not provide memory performance guarantees.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
  \caption{DNN model performance on different in-order
ARM Cortex cores. \fixme{this may be better in the plaform comparison section}.}
  \label{fig:a53_vs_a7}
\end{figure}

%% We do find the performance of the model in the presence of 
%% write co-runners surprising, as it has been previously shown in experiments
%% on another in-order ARM core, the Cortex A7, that write interference was not as 
%% significant\cite{Valsan2016}. We also tested the model, in the presence of write 
%% co-runners, on the A7 cores of the Odroid XU4 platform to see if  the same model 
%% behavior would also be seen there. The results can be seen in Figure ~\ref{fig:a53_vs_a7}. 
%% As expected, when the model is run on A7 cores, the intereference of the write
%% co-runners is not as impactful on model performance as it can still operate at 
%% under 10 Hz.
